---
title: "Comparison of different contour methods"
author: "David J. Hunter"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(tidycensus)
library(sf)
```


## Load data from files

These files can be generated using the scripts `get_acs_tractdata.R` and `countyStats.R`.

```{r}
all_block_geom <- readRDS("data/all_block_geom.rds") 
county_income <- readRDS("data/county_income.rds") 
```

## Read stored data

```{r, warning=FALSE}
fips <- "06071" # san bernardino
fips <- "36061" # New York
fips <- "35001" # bernalillo
    cty <- 
      all_block_geom %>% filter(substr(GEOID,1,5) == fips)
    ctyMedian <- 
      county_income %>% filter(GEOID == fips) %>% .$B19013_001
    cty %>% mutate(
      centroid = suppressWarnings(st_centroid(geometry)),
      X = do.call(rbind, centroid)[, 1],
      Y = do.call(rbind, centroid)[, 2],
      aboveMed = (estimate > ctyMedian)
    ) ->
      cty
    cty %>% dplyr::select(X, Y, aboveMed) %>%
      filter(!is.na(aboveMed)) ->
      allObs
    st_geometry(allObs) <- NULL
    allObs %>% filter(aboveMed) -> success
```

## Plot all block groups

```{r}
cty %>% ggplot(aes(fill=aboveMed)) + geom_sf()
```

## Plot the centroids

```{r}
cty %>% ggplot() + geom_sf() + geom_point(aes(color=aboveMed,x=X, y=Y))
```

## TODO: Compute contours; plot contours

We assume that we have a binary variable (successes/failures) measured on points in the $xy$-plane.

Let $\hat{a}(x,y)$ be the two-dimensional kernel density estimate for all observations (e.g., all census block groups for which we have data). Let $\hat{s}(x,y)$ be the two-dimensional kernel density estimate for all successes (e.g., households with incomes above the median). By Bayes' theorem, we can estimate the conditional probability $f(x,y)$ of a success, given that the observation is taken at point $(x,y)$. This estimate is given by
$$
\hat{f}(x,y) = \frac{\hat{p}\cdot \hat{s}(x,y)}{\hat{a}(x,y)}\text{,}
$$
where $\hat{p}$ is the proportion of observations that are successes. The 50% contour of $\hat{f}(x,y)$ will then be the border between a "success" zone and a "failure" zone.

The `grad()` function computes the gradient of surface `d` at point `(x,y)`.

```{r}
source("scripts/grad.R")
```

## Use `kde2d`for the kernel density estimate. 

```{r}
library(MASS) # masks select
fudge <- 1.0 # fudge factor
abwx <- bandwidth.nrd(allObs$X)*fudge
abwy <- bandwidth.nrd(allObs$Y)*fudge
sbwx <- bandwidth.nrd(success$X)*fudge
sbwy <- bandwidth.nrd(success$Y)*fudge
    s.hat <- kde2d(success$X, success$Y, h = c(sbwx, sbwy), n = 200, lims = c(range(allObs$X),range(allObs$Y)))
    a.hat <- kde2d(allObs$X, allObs$Y, h = c(abwx, abwy), n = 200, lims = c(range(allObs$X),range(allObs$Y)))
    f.hat <- s.hat
#    f.hat$z[f.hat$z < 0.0001] <- 0 # truncate the tails
    f.hat$z <- f.hat$z / a.hat$z * nrow(success) / nrow(allObs)
    f.hat$z[is.nan(f.hat$z)] <- 0 ## replace all 0/0's with 0's
    border <- contourLines(f.hat, levels = 0.5)
borderdf <- list()
for(i in 1:length(border)){
  borderdf[[i]] <- data.frame(x = border[[i]]$x, y = border[[i]]$y)
}

cat("Found", length(border), "contours.\n")
```

Plot the contours.

```{r}
cty %>% ggplot() + geom_sf() + geom_point(aes(color=aboveMed,x=X, y=Y)) -> p
for(i in seq(length(border))) {
  p <- p + geom_path(data=borderdf[[i]], aes(x=x, y=y), color="red3") 
}
p
```

## Use `ash`for the kernel density estimate. 

```{r}
library(ash)
aspect <- (max(allObs$X) - min(allObs$X))/(max(allObs$Y) - min(allObs$Y))
    nbin = c(round(500*aspect), 500)
    m = c(round(20*aspect),20)
    allRange <-
      matrix(c(range(allObs$X), range(allObs$Y)), 2, 2, byrow = TRUE)
    xa <- data.matrix(allObs[, 1:2])
    xs <- data.matrix(success[, 1:2])
    binsa <- bin2(xa, allRange, nbin)
    binss <- bin2(xs, allRange, nbin)
    a.hat <- ash2(binsa, m)
    s.hat <- ash2(binss, m)
    f.hat <- s.hat
    f.hat$z <- f.hat$z / a.hat$z * nrow(success) / nrow(allObs)
    f.hat$z[is.nan(f.hat$z)] <- 0 ## replace all 0/0's with 0's
    border <- contourLines(f.hat, levels = 0.5)
borderdf <- list()
for(i in 1:length(border)){
  borderdf[[i]] <- data.frame(x = border[[i]]$x, y = border[[i]]$y)
}

cat("Found", length(border), "contours.\n")
```

Plot the contours.

```{r}
cty %>% ggplot() + geom_sf() + geom_point(aes(color=aboveMed,x=X, y=Y)) -> p
for(i in seq(length(border))) {
  p <- p + geom_path(data=borderdf[[i]], aes(x=x, y=y), color="red3", size=1.5) 
}
p
```


## Use `KernSmooth`for the kernel density estimate. 

```{r}
library(KernSmooth)
se <- "minim" # estimate of scale
knl <- "normal" # smoothing kernel
lv <- 3 # levels in plug-in rule
gs <- 100 # grid size
fudge <- 1.0 # fudge factor
abwx <- dpik(allObs$X, scalest = se, kernel = knl, gridsize = gs, level = lv) * fudge
abwy <- dpik(allObs$Y, scalest = se, kernel = knl, gridsize = gs, level = lv) * fudge
sbwx <- dpik(success$X, scalest = se, kernel = knl, gridsize = gs, level = lv) * fudge
sbwy <- dpik(success$Y, scalest = se, kernel = knl, gridsize = gs, level = lv) * fudge
    a.hat <- bkde2D(cbind(allObs$X,allObs$Y), c(abwx, abwy), gridsize = c(gs, gs))
    names(a.hat) <- c("x", "y", "z") # to match other functions
    a.hat$z[a.hat$z < 0.05] <- 0 # truncate the tails
    s.hat <- bkde2D(cbind(success$X,success$Y), c(sbwx, sbwy), gridsize = c(gs, gs))
    names(s.hat) <- c("x", "y", "z") # to match other functions
    f.hat <- s.hat
    f.hat$z <- f.hat$z / a.hat$z * nrow(success) / nrow(allObs)
    f.hat$z[is.nan(f.hat$z)] <- 0 ## replace all 0/0's with 0's
    border <- contourLines(f.hat, levels = 0.5)
borderdf <- list()
for(i in 1:length(border)){
  borderdf[[i]] <- data.frame(x = border[[i]]$x, y = border[[i]]$y)
}

cat("Found", length(border), "contours.\n")
```

Plot the contours.

```{r}
cty %>% ggplot() + geom_sf() + geom_point(aes(color=aboveMed,x=X, y=Y)) -> p
for(i in seq(length(border))) {
  p <- p + geom_path(data=borderdf[[i]], aes(x=x, y=y), color="red3", size=1.5) 
}
p
```

## Try the `ks` package


```{r}
library(ks)
a.hat <- kde()
```


